% ----------------------------------------------
%
%	Damian Skrzypiec
% 	02.11.2017	
% 	This script contains overview of impelementation
% 	of cglearn package.	
%
% ----------------------------------------------


In this project we also provide an implementation of LCD algorithm. We make the implementation available in R through package called \textit{cglearn}.
R is broadly used, both in academia and in business, open-source statistical programming language. The name of package is short for chain graph structure learning.
It is also analogy to the R package \textit{bnlearn} for learning bayesian networks.
The \textit{cglearn} package could be installed through Github and soon will be available on CRAN. Details about package installation are in subsection \ref{subsec:PackageInstall}.

The package provides functionality for learning chain graphs structure within LCD algorithm and performe conditional independence analysis for given 
data set. The functionality is further described in subsection \ref{subsec:PackageFunc}.

It is worth to mention that current implementation of LCD algorithm provides functionality of learning chain graphs only for numeric categorical random variables.
The major part of package was implemented in C++ and injected in R using Rcpp package. It was done in purpose of maximizing performance of algorithm. 
The main part of LCD algorithm implementation took over two thousands lines of code. The source code can be found on GitHub repository at \cite{GIT}.











