% ----------------------------------------------
%
%	Damian Skrzypiec
% 	02.11.2017	
% 	This script contains short description
% 	cglearn package functionality.	
%
% ----------------------------------------------


In this section we describe the most important functionalities of \textit{cglearn} package.
In the following subsections we will give short description of the package's function, definition
of arguments and the result. At the end of this section we suggest possible further improvements of 
the \textit{cglearn} package.


\subsection{cglearnLCD}
The function \textit{cglearnLCD} is the main function of the package. It performes LCD algorithm for the 
following given arguments.

\begin{itemize}
	\item \textbf{data} - \textit{data.frame} object with categorical numeric variables.

	\item \textbf{separationTree} - \textit{list} of vectors column ids which defines separation tree based on given data.

	\item \textbf{pValueThr} - \textit{numeric} number from interval $(0, 1)$ which represents p-value threshold to determine conditional independence.
\end{itemize}

The \textit{cglearnLCD} function returns a \textit{data.frame} object which represents incidence matrix of chaing graph obtained by LCD algorithm.
 

\subsection{calcCondIndepend}
The function \textit{calcCondIndepend} performs statistical test for conditional independence $x \bigCI y \mid Z$. Required arguments
are listed below.

\begin{itemize}
	\item \textbf{x} - \textit{numeric} categorical vector.

	\item \textbf{y} - \textit{numeric} categorical vector.

	\item \textbf{listOfNumVect} - \textit{list} of categorical numeric vectors.
\end{itemize}

The \textit{calcCondIndepend} function returns a S3 object which contains value of statistic, number of degrees of freedom and p-value of the test.





