% ----------------------------------------------
%
%	Damian Skrzypiec
% 	August 2017
%	Skeleton recovery algorithm.
%
% ----------------------------------------------


Skeleton Recovery algorithm is composed of three parts. In the first part of this algorithm (lines 3-11) local skeletons are recovered.
This is obtained by looping over all nodes in $ \mc{T}(G, \mc{C})$. For given node $C_h \in \mc{T}(G, \mc{C})$ we create
complite graph $G_h = (C_h, E_h)$ and we test conditional independence of all possible pairs of $C_h$. For fixed pair
$\{u, v\} \in C_h$ we every subset $S_{uv} \subset C_h$ we test condition $u \bigCI v \mid S_{uv}$. If such a condition
is satisfied then edge $(u, v)$ is removed from local graph $G_h$. By condition $1$ from theorem \ref{LCDSkeletonThm} we observe
that edge which is removed from local skeleton cannot occure in global skeleton. This observation increase performance
of implementation of Skeleton Recovery algorithm. In the second part of Skeleton Recovery algorithm (lines 12-17) local
skeletons are combined into global skeleton and some of extra edges are removed based on second condition in theorem \ref{LCDSkeletonThm}.
The third part of this algorithm (lines 18-24) also focus on removing incorrect edges. This part of algorithm are backed up by
third condition of theorem \ref{LCDSkeletonThm}. The following algorithm was introduced in \cite{CG}, chapter 3.2, algorithm 1.

\begin{algorithm}
	\caption{(LCD) Skeleton Recovery}\label{skeletonRecoveryAlg}
	
	\textbf{Input:} A separation tree $ \mc{T}(G, \mc{C})$; 
					perfect conditional independence knowledge about $\mathbb{P}$.  \\
	\textbf{Output:} The skeleton $G^{'}$ of $G$; a set $\mc{S}$ of c-separators.

	
	\begin{algorithmic}[1]
		\Procedure{RecoverySkeleton($ \mc{T}(G, \mc{C})$)}{}
			\State $\mc{S} = \emptyset$
	
			\ForAll{$\mbox{node} \ C_h \in \mc{T}(G, \mc{C})$} 
				\State	Create complete undirected graph $G_h = (C_h, E_h)$;
				\ForAll{$\mbox{vertex pair} \ \{u, v \} \subset C_h $}
					\If{$\exists S_{uv} \subset C_h \ u \bigCI v \mid S_{uv}$}
						\State Delete edge $(u, v)$ from graph $G_h$;
						\State $\mc{S} := \mc{S} \cup S_{uv}$; \Comment{Add set $S_{uv}$ to separators}
					\EndIf
				\EndFor
			\EndFor			
		
			\State Combine all the graphs $(G_h)_{i \in \{1, \dots, H}$ into undirected graph 
			$G^{'} = (V, \bigcup_{h = 1}^{H} E_h)$;
			
			\ForAll{$ \{u, v \} \in G^{'} \ \mbox{contained in more then 
					one node of} \ \mc{T}(G, \mc{C})$}
				\If{$ \exists C_h \ \{u, v \} \subset C_h \ \mbox{and} \ (u, v) \notin E_h$}
					\State Delete the edge $(u, v)$ from $G^{'}$;
				\EndIf
			\EndFor
			
			\ForAll{$ \{u, v \} \in G^{'} \ \mbox{contained in more then 
					one node of} \ \mc{T}(G, \mc{C})$}
					
				\State $N_{uv} := \{S \subset \mbox{ne}_{G^{'}}(u) \cup \mbox{ne}_{G^{'}}(v) \ | \ 
									S \not\subset C_h \ \mbox{and} \ \{u, v\} \subset C_h  \}$
									
				\If{$u \bigCI v \mid S_{uv} \ \mbox{for some} \ S_{uv} \subset N_{uv}$}
					\State Delete edge $(u, v)$ from graph $G^{'}$;
					\State $\mc{S} := \mc{S} \cup S_{uv}$; \Comment{Add set $S_{uv}$ to separators}
				\EndIf
			\EndFor	
			
			\State \textbf{return:} $G^{'}$, $\mc{S}$.
		\EndProcedure
	\end{algorithmic}
\end{algorithm}


% -----------------------------------
%	Example of application LCD alg.
% -----------------------------------

\begin{ex} \label{skeletonRecoveryEx}
	To illustrate execution of the LCD algorithm let's assume we have data which joint distribution is faithful to
	graph presented in figure \ref{fig:graphForNodeTree} but we do not know it's chain graph representation yet.
	Additionally we have separation tree presented on figure \ref{fig:nodeTree}. 
	Result of first phase of skeleton recovery algorithm is presented on figure \ref{fig:LCDfirstPhaseResult}. 
	For every node tree we have local undirected graph representing local
	(in sense of particular node) independence structure. Phase two of skeleton recovery algorithm join local graphs into
	global undirected graph and removes some of redundant edges. 
	Result of execution second phase is represented on figure
	\ref{fig:LCDsecondPhaseResult}. Result of applying skeleton recovery algorithm is the same as outcome from second
	phase of the algorithm, because there isn't pair of random variable satisfying condition from 20th line of Algorithm 
	\ref{skeletonRecoveryAlg}. Output set of separators in this example is $\mc{S} = \{ \{B, C \}, \{F \} \}$.

\end{ex}

% -----------------------------------
%	Viz. of I phase of LCD alg.
% -----------------------------------
\begin{figure}
	\centering
	\vspace{-10pt}
	\begin{tikzpicture}[>=stealth',shorten >=1pt,auto,node distance=2.5cm,
	                    thick,main node/.style={circle,draw,font=\sffamily\Large\bfseries}]
		% Nodes
		  %Row1
		  \node (A1) at (0, 3) [main node] {A};
		  \node (I1) at (2, 3) [main node] {I};
		  
		  \node (A2) at (4, 2) [main node] {A};
		  \node (B2) at (4, 4) [main node] {B};
		  \node (C2) at (6, 4) [main node] {C};
		  \node (D2) at (6, 2) [main node] {D};

		  \node (D3) at (8, 4) [main node] {D};
		  \node (E3) at (8, 2) [main node] {E};
		  \node (F3) at (10, 4) [main node] {F};
		  
		  % Row2
		  \node(E4) at (0, -1) [main node] {E};
		  \node(F4) at (2, -1) [main node] {F};
		  \node(H4) at (4, -1) [main node] {H};
		  
		  \node(D5) at (6, -2) [main node] {D};
		  \node(F5) at (8, -2) [main node] {F};
		  \node(G5) at (8, 0) [main node] {G};			
		
		% Edges
		  \draw (A1) -- (I1);
		  \draw (A2) -- (B2) -- (C2) -- (D2);
		  \draw (E3) -- (D3) -- (F3) -- (E3);			  
		  
		  \draw (E4) -- (F4) -- (H4);
		  \draw (D5) -- (F5) -- (G5);
		  
		\end{tikzpicture}
		
	\caption{Result of first phase of LCD algorithm}			
	\label{fig:LCDfirstPhaseResult} 
\end{figure}




% -----------------------------------
%	Viz. of II phase of LCD Alg
% -----------------------------------
\begin{figure}
		\centering
		\vspace{-10pt}
		\begin{tikzpicture}[>=stealth',shorten >=1pt,auto,node distance=2.5cm,
		                    thick,main node/.style={circle,draw,font=\sffamily\Large\bfseries}]
			% Nodes
			  \node[main node] (A) {A};
			  \node[main node] (B) [above of = A] {B};
			  \node[main node] (C) [right of = B] {C};
			  \node[main node] (D) [below of = C] {D};
			  \node[main node] (E) [below of = D] {E};
			  \node[main node] (F) [right of = D] {F};
			  \node[main node] (G) [above of = F] {G};
			  \node[main node] (H) [below of = F] {H};			
			  \node[main node] (I) [below of = A] {I};			
			
			% Edges
			  \draw (I) -- (A);
			  \draw (A) -- (B);
			  \draw (B) -- (C);
			  \draw (C) -- (D);
			  \draw (D) -- (F) -- (E);
			  \draw	(D) -- (E);
			  \draw (F) -- (G);
			  \draw (F) -- (H);
			\end{tikzpicture}
			
		\caption{Result of second phase of LCD algorithm}			
		\label{fig:LCDsecondPhaseResult} 
	\end{figure}
